---
title: "Project memo"
author: "Team name"
output: github_document
editor_options: 
  chunk_output_type: console
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
library(stringr)
library(dplyr)
library(readxl)
library(ggstream)
library(showtext)
library(ggtext)
library(gghighlight)
options(scipen = 999)
```

## Data Clean Up Steps for Overall Data

### Step 1: Load datasets

#### Step 1a: Load fossil fuel data and change column names

``` {r clean_data, message = FALSE, warning = FALSE}
fuels <- read_csv("../data/Dead-River.csv",
                  col_names = c("X1", "X2", 
                                "Delivery_date", "Fuel_type",
                                "Tank_number", "Building",
                                "Gallons", "Unit_cost",
                                "Cost"))
```

#### Step 1b: Load building area data
```{r}
area <- read_csv("../data/BuildingArea.csv",
                 col_names = c("Building", "Square_feet"))
```


#### Step 1c: Join area data to fuels data

```{r}
fuels <- fuels |>
left_join(area)
```


### Step 2: Change Delivery Date & Remove top extra row
```{r change_delivery_date_and_remove_top_row}
fuels <- fuels |>
select(Delivery_date, Fuel_type, Tank_number, Building, Gallons, Unit_cost, Cost, Square_feet) |>
  filter(Delivery_date != "Delivery Date") |> 
  mutate(Delivery_date = mdy(Delivery_date))
 
```


### Step 3: Change tank number to integer values
```{r change_tank_number}
fuels <- fuels |>  
  mutate(Tank_number = case_when(
   Tank_number == "Tank 8" ~ "8",
   Tank_number == "Tank 3" ~ "3",
   Tank_number == "Tank 24" ~ "24",
   Tank_number == "Tank 17" ~ "17",
   TRUE ~ Tank_number
  ),
  Tank_number = as.integer(Tank_number))
```

### Step 4: Change cost to numeric values

```{r change_cost}
fuels <- fuels |>
  mutate(Cost = case_when(
    Cost == "610.12 cr" ~ "-610.12", # the cr denotes a credit/refund, so here it is changed to a negative numeric value
    TRUE ~ Cost
  ),
  Cost = as.numeric(Cost))
```

### Step 5: Standardize fuel type names
```{r standardize_fuel_type}
fuels <- fuels |>  
  mutate(Fuel_type = case_when(
    Fuel_type == "#2 HEATING OIL" ~ "Heating Oil",
    Fuel_type == "#2 Heating Oil" ~ "Heating Oil",
    Fuel_type == "SLPP #2 FUEL OIL" ~ "Heating Oil",
    Fuel_type == "SLP LIQ PROPANE" ~ "Propane",
    Fuel_type == "LIQUID PROPANE" ~ "Propane",
    Fuel_type == "DYED KEROSENE" ~ "Dyed Kerosene",
  TRUE ~ Fuel_type
  ))
```

### Step 6: Change gallons and unit cost to numeric values

```{r change_gallons_and_unit_cost}
fuels <- fuels |>
  mutate(Gallons = as.numeric(Gallons)) |>
  mutate(Unit_cost = as.numeric(Unit_cost))
```


### Step 7: Add year and month variables
```{r add_year_month}
fuels <- fuels |>
  mutate(Year = substr(Delivery_date, 1, 4)) |>
  mutate(Month = substr(Delivery_date, 6, 7)) |>
  mutate(Year = as.numeric(Year)) |>
  mutate(Month = as.numeric(Month))
```

### Step 8: Change square feet to numeric

```{r change_to_numeric}
fuels <- fuels |>
  mutate(Square_feet = case_when(
    Building == "Turrets" ~ "8400",
    Building == "Turrets Annex" ~ "4200",
    TRUE ~ Square_feet
  )) |>
  mutate(Square_feet = as.numeric(Square_feet))
```


### Step ?: Ignore for now I'm working on it but am confused

```{r pie-data}

# make one big function



# main_total_fuels |>
#   filter(Building == c("Witchcliff Apartments", "Witchcliff","Studio 5+6","PRF",
#                        "Pottery Studio","Peggy Barn", "Peach House", "Hatchery", "Greenhouse", 
#                        "Davis Carriage", "Cottage", "CHE Generator", "Carriage", 
#                        "BHF New Greenhouse", "BHF Main Bldg/2 Greenhouse", "BHF Farm House", 
#                        "18B Norris Ave", "171 Beech Hill Road", "14 Norris Ave")) |>
#   (Other_Gallons <- sum(Total_Gallons))

 # main_total_fuels <- data.frame(
 #  Building = unique(fuels$Building),
 #  Total_Gallons = unique(fuels$Total_Gallons_per_Building)
 # )

 # with(main_total_fuels, sum(Total_Gallons[Total_Gallons < 9952.2]))

#  main_total_fuels <- main_total_fuels |>
#    add_row(Building = "Other",
#           Total_Gallons = with(main_total_fuels, sum(Total_Gallons[Total_Gallons < 41871.1])))
# 
# main_total_fuels <- main_total_fuels |>
#   filter(Total_Gallons >= 41871.1 | Building == "Other")


```




## Plots

### ggsave example for saving plots

```{r starwars ggsave example}
p1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```



### Plot 1: Fuel Prices Over Time

#### Final Plot 1

```{r prices_over_time}
fuels %>% 
  ggplot(mapping = aes(x = Delivery_date, y = Unit_cost, color = Fuel_type)) +
  geom_line() +
  facet_wrap(~ Fuel_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  guides(color = FALSE) +
  labs(title = "Fuel Prices Over Time",
       subtitle = "2014 - 2024",
       x = "Delivery date",
       y = "Cost per gallon in USD")

```

### Plot 2: Total Amount of Fuel per Building

```{r total_fuel_per_building, fig.alt = "Horizontal bar graph showing the building on the y-axis and the total amount of gallons of fuel used by those buildings between 2014 and 2024. The fill color of the graph indicates whether the building uses heating oil or propane. The graph shows the 5 buildings that consumed the most amount of fuel in this time span: Arts & Science + Gates, Kaelber, Blair Tyson, Dorr Natural History Museum and Seafox. The graph also shows that Arts & Science + Gates, the biggest consumer with more than 150000 gallons of fuel, uses propane, while the other 4 buildings use heating oil. Seafox, the smallest consumer out of the 5 used around 45000 gallons."}

fuels |>
  group_by(Building, Fuel_type) |>
  summarize(Total_Gallons_per_Building = sum(Gallons), .groups = "drop") |>
  slice_max(order_by = Total_Gallons_per_Building, n = 5) |>
  slice_max(Total_Gallons_per_Building, n = 5) |>
  ggplot(aes(y = fct_reorder(Building, Total_Gallons_per_Building), x = Total_Gallons_per_Building, fill = Fuel_type)) +
  geom_col() +
  theme_minimal() +
  scale_fill_viridis_d() +
  guides(color = FALSE) +
  labs(title = "Buildings Using the Most Amount of Fuel",
       subtitle = "2014 - 2024",
       y = "Building",
       x = "Total Amount of Fuel in Gallons",
       fill = "Fuel Type")

ggsave("Buildings-Using-Most-Fuel.png")

```

### Plot 3: 5 biggest over time

```{r 5_biggest_over_time}
fuels |>
  filter(Building %in% c("Blair Tyson", "Arts & Sci + Gates", "Kaelber"), Year < 2024) |>
  group_by(Building, Year) |>
  summarize(year_sum = sum(Gallons), .groups = "drop") |>
  ggplot(aes(x = Year, y = year_sum, color = Building, group = Building)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Total Gallons of Fuel per Year",
       subtitle = "2014 - 2023, 3 Biggest Consumers",
       y = "Gallons",
       caption = "HPWH = electric heat pump water heating system") +
  scale_x_continuous(breaks = seq(from = 2014, to = 2024, by = 1)) +
  ylim(0, NA) +
  scale_color_viridis_d() +
  geom_vline(xintercept = 2020, linetype = "dotted", colour = "blue") +
  geom_text(aes(x = 2020.5, y = 16000), label = "COVID", colour = "blue") +
  geom_text(aes(x = 2021.3, y = 8300), label = "BT HPWH installed", colour = "black", size = 3) +
  geom_point(aes(x = 2022, y = 7500), color = "black", size = 3, shape = 15) +
  geom_line(aes(y = 900), color = "darkgreen") +
  geom_text(aes(x = 2021.75, y = 1500), label = "Well insulated family house", color = "darkgreen", size = 3)

ggsave("3-biggest-consumers-over-time.png", width = 8, height = 4)
```

### Plot 4: Fuel usage per month
```{r fuel_per_month}
fuels |> 
  filter(Year < 2024) |>
  group_by(Month, Fuel_type) |>
  summarize(month_sum = sum(Gallons), .groups = "drop") |>
  ggplot(aes(x = Month, y = month_sum, fill = Fuel_type)) +
  geom_col() +
  labs(title = "Fuel Usage per Month",
       subtitle = "2014 - 2023, all campus-owned buildings",
       y = "Total Gallons",
       fill = "Fuel Type") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  scale_fill_viridis_d()

ggsave("fuel_usage_per_month.png", width = 8, height = 4)
```

### Plot ?: A bunch that I'll rename and clean up later 

```{r gallons_per_fuel_type}
fuels |>
  group_by(Fuel_type) |>
  mutate(Total_Gallons_per_Fuel_Type = sum(Gallons)) |>
ggplot(aes(x = Fuel_type, y = Total_Gallons_per_Fuel_Type)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_minimal() +
  guides(color = FALSE) +
  labs(title = "Total Amount of Fuel per Fuel Type",
       x = "Fuel Type",
       y = "Amount of Fuel")
```

```{r gallons_per_year}
fuels |>
  group_by(Fuel_type) |>
  mutate(Total_Gallons_per_Fuel_Type = sum(Gallons)) |>
  filter(Year != 2024) |>
  ggplot(aes(x = Year, y = Total_Gallons_per_Fuel_Type)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_minimal() +
  guides(color = FALSE) +
  labs(title = "Total Amount of Fuel per Year",
       x = "Year",
       y = "Amount of Fuel")
```

```{r gallons_per_fuel_type_per_year_facet_buildings}

fuels |>
  filter(Year != 2024) |>
  group_by(Year, Building) |>
  mutate(Total_Gallons_per_years = sum(Gallons)) %>%
  ggplot(aes(x = Year, y = Total_Gallons_per_years, fill = Fuel_type))+
  geom_col() +
  facet_wrap(~Building)
  theme_minimal()

```

```{r testing}

fuels |>
  filter(Building == "Kaelber" & Year != 2024) |>
  group_by(Year) |>
  #View() |>
  mutate(Total_Gallons_per_Year = sum(Gallons)) |>
  ggplot(aes(x = Year, y = Total_Gallons_per_Year)) +
  geom_col()

```

```{r pie-total-gallons}

# do factor shift thingy

#  ggplot(main_total_fuels, aes(x = "", y = Total_Gallons, fill = Building)) +
#   geom_bar(stat = "identity", width = 1, color = "white") +
#   coord_polar("y", start = 0) +
#   theme_void()
# 
# ggplot(main_total_fuels, aes(y = Building, x = Total_Gallons)) +
#   geom_col()

```


### Plot ?: Building area
```{r building_area}
fuels |>
  filter(Square_feet >= 5000) |>
  ggplot(aes(x = Building, y = Square_feet)) +
  geom_col()
```

### Plot ?: Gallons per square foot

```{r total_gallons_per_sf}
fuels |>
  filter(Year >= 2019, Year <= 2023) |>
  group_by(Building) |>
  mutate(Total_Gallons_per_Building = sum(Gallons)) |>
  mutate(total_per_sf = (Total_Gallons_per_Building / Square_feet)) |>
  summarize(total_per_sf = unique(total_per_sf)) |>
  filter(total_per_sf > 2) |>
  ggplot(aes(x = Building, y = total_per_sf)) +
  geom_col() +
  labs(title = "Total Gallons per Square Foot",
       subtitle = "Highest 6 Buildings",
       y = "Total Gallons per Square Foot") +
  theme_minimal() +
  coord_flip()
```

### Plot showing gallons per month before and after heat pump - Not useful in the end

```{r before-and-after-heat-pump}

fuels |>
  filter(Building == "Kaelber" & Year >= 2019 & Year <= 2023) |>
  group_by(Year, Month) |>
  mutate(Total_Gallons_per_Month = sum(Gallons)) |>
  #View() |>
  group_by(Year) |>
  ggplot(aes(x = Month, y = Total_Gallons_per_Month, group = Year, colour = Year)) +
  geom_line() +
  geom_point() +
  facet_wrap(~Year)

```

```{r fuel-type-per-month}

fuels|>
  filter(Year != 2024) |>
  group_by(Year, Fuel_type) |>
  mutate(Total_Gallons_per_Year = sum(Gallons)) |>
  ggplot(aes(x = Year, y = Total_Gallons_per_Year, colour = Fuel_type)) +
  geom_line() +
  facet_wrap(~Fuel_type) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Plot ?: Fuel usage per month
```{r fuel_per_month_unused}
fuels %>% 
  ggplot(mapping = aes(x = Month, y = Gallons, color = Fuel_type, group = Fuel_type)) +
  facet_wrap(~ Fuel_type) +
  geom_point() +
  geom_line() +
  scale_color_viridis_d() +
  theme_minimal() +
  guides(color = FALSE) +
  labs(title = "Fuel Usage by Month",
       x = "Month",
       y = "Gallons")
```

```{r plot}

pal=c("#003f5c",
      "#2f4b7c",
      "#665191",
      "#a05195",
      "#d45087",
      "#f95d6a",
      "#ff7c43",
      "#ffa600")

order <- c("Seafox", "Dorr NHM", "Blair Tyson", "Kaelber", "Arts & Sci + Gates")

# 
# 
# test <- fuels |>
#   filter(Building %in% c("Blair Tyson", "Arts & Sci + Gates", "Kaelber", "Seafox", "Dorr NHM"), 
#          Year != 2024) |>
#   group_by(Year, Building) |>
#   mutate(Total_Gallons_per_years = sum(Gallons))
# 
# length(test) <- length(fuels)
# 
# class(test$Building)

fuels |>
  filter(#Building %in% c("Blair Tyson", "Arts & Sci + Gates", "Kaelber", "Seafox", "Dorr NHM"), 
         Year != 2024) |>
  group_by(Year, Building) |>
  summarize(Total_Gallons_per_years = sum(Gallons)) |>
  filter(Total_Gallons_per_years != "NA") %>% 
  #mutate(Building = factor(Building, levels=order)) |>
  ggplot(aes(Year, Total_Gallons_per_years, fill = Building)) +
  geom_area() +
  
  scale_fill_manual(values=pal) +
  scale_color_manual(values = pal) +
  gghighlight(Building %in% 
                c("Blair Tyson", "Arts & Sci + Gates", "Kaelber", "Seafox", "Dorr NHM")) +
  
  theme(
    axis.line.x = element_line(linewidth = .75),
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(color="black", size=10),
    plot.margin = margin(10,50,10,10),
    panel.background = element_rect(fill = "white", colour = "white", 
                                   linetype = "solid", linewidth = 0.5),
    legend.position = "none",
  ) +
  labs(title = "Progression of the Top Five Fuel Consumers",
       subtitle = "2014 - 2023",
       y = "",
       x = "Year") +
  
# Labels
  annotate("text", x = 2023.1, y = 32000,
           label = "Seafox",
           hjust=0,
           size=3,
           lineheight=.8,
           fontface="bold",
           color=pal[1]) +
  
  annotate("text", x = 2023.1, y = 27500,
           label = "Dorr NHM",
           hjust=0,
           size=3,
           lineheight=.8,
           fontface="bold",
           color=pal[2]) +
  
  annotate("text", x = 2023.1, y = 23000,
           label = "Blair Tyson",
           hjust=0,
           size=3,
           lineheight=.8,
           fontface="bold",
           color=pal[3]) +
  
  annotate("text", x = 2023.1, y = 18000,
           label = "Kaelber",
           hjust=0,
           size=3,
           lineheight=.8,
           fontface="bold",
           color=pal[4]) +
  
  annotate("text", x = 2023.1, y = 10000,
           label = "Arts & Sci + Gates",
           hjust=0,
           size=3,
           lineheight=.8,
           fontface="bold",
           color=pal[5]) +
  
  annotate("text", x = 2025, y = 27500,
           label = "",
           hjust=0,
           size=3,
           lineheight=.8,
           fontface="bold",
           color=pal[2])

ggsave("top-5-fuel-area-plot.png", width = 4, height = 4)


```

```{r}


fuels |>
  filter(Year > 2022) |>
  group_by(Building) |>
  mutate(Total_Gallons_per_Buildings = sum(Gallons)) |>
  ggplot(aes(x = Total_Gallons_per_Buildings, y =  fct_reorder(Building, Total_Gallons_per_Buildings, .fun = sum), fill = Building)) +
  geom_col() +
  gghighlight(Building %in% 
                c("Blair Tyson", "Arts & Sci + Gates", "Kaelber", "Seafox", "Dorr NHM")) +
  
  scale_fill_manual(values = pal) +
  scale_color_manual(values = pal) +
  
  theme(
    axis.line.x = element_line(linewidth = .75),
    panel.grid = element_blank(),
    axis.text.x = element_text(color="black", size=10),
    panel.background = element_rect(fill = "white", colour = "white", 
                                   linetype = "solid", linewidth = 0.5),
    legend.position = "none",
  ) +
  labs(title = "Use of Fuel Across COA Buildings",
       subtitle = "2022 - 2024",
       y = "",
       x = "Total Amount of Fuel in Gallons")

ggsave("top-5-fuel-bar-plot.png", width = 4, height = 4)


```

```{r}


fuels |>
  group_by(Year, Building) |>
  mutate(Total_Gallons_per_years = sum(Gallons)) |>
  mutate(fuel_per_sf = (Total_Gallons_per_years/Square_feet)) |>
  filter(Building %in% c("B&G", "Davis Centre", "Dorr NHM", "Hatchery", "Seafox")) |>

  # summarize(total_per_sf = unique(total_per_sf)) |>
  ggplot(aes(Year, fuel_per_sf, fill = Building, label = Building, color = Building)) +
  geom_area() +
  
  scale_fill_manual(values=pal) +
  scale_color_manual(values = pal) +
  
  theme(
    axis.line.x = element_line(linewidth = .75),
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(color="black", size=10),
    panel.background = element_rect(fill = "white", colour = "white", 
                                   linetype = "solid", linewidth = 0.5),
    legend.position = "none",
  )

```