---
title: "Project memo"
author: "Team name"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
```

## Data Clean Up Steps for Overall Data

### Step 1: _________

``` {r clean_data, message = FALSE, warning = FALSE}
fuels <- read_csv("../data/Dead-River.csv",
                  col_names = c("X1", "X2", 
                                "Delivery_date", "Fuel_type",
                                "Tank_number", "Building",
                                "Gallons", "Unit_cost",
                                "Cost")) |>
  select(Delivery_date, Fuel_type, Tank_number, Building, Gallons, Unit_cost, Cost) |>
  filter(Delivery_date != "Delivery Date") |> 
  mutate(Delivery_date = mdy(Delivery_date)) |>
  mutate(Tank_number = case_when(
  Tank_number == "Tank 8" ~ "8",
  Tank_number == "Tank 3" ~ "3",
  Tank_number == "Tank 24" ~ "24",
  Tank_number == "Tank 17" ~ "17",
  TRUE ~ Tank_number
),
Tank_number = as.integer(Tank_number)) %>% 
mutate(Cost = case_when(
    Cost == "610.12 cr" ~ "-610.12",
    TRUE ~ Cost
  ),
  Cost = as.numeric(Cost)) %>% 
  mutate(Gallons = as.numeric(Gallons)) %>% 
  mutate(Unit_cost = as.numeric(Unit_cost)) %>% 
  mutate(Fuel_type = case_when(
    Fuel_type == "#2 HEATING OIL" ~ "Heating Oil",
    Fuel_type == "#2 Heating Oil" ~ "Heating Oil",
    Fuel_type == "SLPP #2 FUEL OIL" ~ "Heating Oil",
    Fuel_type == "SLP LIQ PROPANE" ~ "Propane",
    Fuel_type == "LIQUID PROPANE" ~ "Propane",
    Fuel_type == "DYED KEROSENE" ~ "Dyed Kerosene",
  TRUE ~ Fuel_type
  ))
```


### Step 2: ________

## Plots

### ggsave example for saving plots

```{r starwars ggsave example}
p1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```



### Plot 1: Fuel Prices Over Time

#### Final Plot 1

```{r}
fuels %>% 
  ggplot(mapping = aes(x = Delivery_date, y = Unit_cost, color = Fuel_type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ Fuel_type) +
  scale_color_viridis_d() +
  theme_minimal() +
  guides(color = FALSE) +
  labs(title = "Fuel Prices Over Time",
       subtitle = "2014 - 2024",
       x = "Delivery date",
       y = "Cost per gallon in USD")

```

### Plot 2: Total Amount of Fuel per Building

```{r}
fuels %>%
ggplot(mapping = aes(x = Building)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_minimal() +
  guides(color = FALSE) +
  labs(title = "Total Amount of Fuel per Building",
       x = "Building",
       y = "Amount of Fuel")

```

### Plot 3: ___________

Add more plot sections as needed. Each project should have at least 3 plots, but talk to me if you have fewer than 3.

### Plot 4: ___________ 

